Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2021-09-06 ms/manuscript.Rmd
Update MS
- Fix mentions of Honshu
- Change github URL

Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because the necessary data (DNA sequences and georeferenced occurrence records) and analytic tools have only become available relatively recently. 
These datasets and tools now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism. 
Better ways to characterize biological regions ("bioregions") are also available [e.g., @Laffan2016; @White2019; @Daru2020a], rather than relying on ad-hoc characterizations.
Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.

@@

A necessary step in any analysis of biodiversity is to set the grain size used to accurately define communities. 
Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. 
We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species). 
At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). 
We calculated sampling redundancy (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. 
Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
We filtered out any cells with redundancy < 0.1 and used this size for subsequent analyses.

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")` [@R-raster], and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`). 
We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.

@@

Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]. 
However, using this dataset alone for phylogeny suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils available for molecular dating. 
Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree. 
To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa). 
We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with 1,000 ultrafast bootstrap replicates [@Minh2013].
Molecular dating was conducted with `r software("treePL")` v1.0 [@Smith2012] using `r n_fossil_points` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020]. 
The root of the tree (most recent common ancestor of all land plants) was fixed to 475 mya [@Testo2016a].
We trimmed the resulting ultrametric tree to only Japanese taxa.
treePL requires branch lengths to be > 0, and it resets extremely small branches an arbitrary minimum positive length.
This may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* dataset.
We converted any clades in the original dated tree consisting of identical sequences to polytomies with zero branch length between OTUs.
This tree was used for all subsequent analysis unless indicated otherwise.

@@

We tested for phylogenetic signal in continuous traits using two alternative metrics, Blomberg's *K* [@Blomberg2003] and Pagel's lambda [@Pagel1999]. 
*K* describes the ratio between the observed variance in a trait vs. the amount of variance expected under Brownian Motion (BM); lambda is a scaling parameter that transforms the phylogeny such that trait values fit those most closely expected under BM. 
Both lambda and *K* = 1 under BM; lower values indicate less signal (overdispersion) and higher values indicate more signal (clustering). 
We measured *K* and lambda using the `r func("phylosig")` function in the R package `r pack("phytools")` [@Revell2012], which also calculates significance based on a randomization test for *K* and a likelihood ratio test for lambda. 
We tested for phylogenetic signal in qualitative (coded as binary) traits using Fritz and Purvis' *D* [@Fritz2010] with the `r func("phylo.d")` function in the R package `r pack("caper")` [@Orme2018], which also assesses significance by comparing the observed value to a null distribution of random values and a simulated distribution of values expected under BM.
These methods for phylogenetic signal cannot use a tree with zero-length polytomies, so we used the original dated tree from treePL.

@@

We constructed a linear mixed model for each biodiversity metric dependent on the environmental variables ("environmental models"). 
Initial inspection of the relationship between each biodiversity metric and temperature showed a hump-shaped pattern, so we also included a quadratic term for temperature only. 
Richness was fit with the negative binomial response family; others were Gaussian. 
We fit models including a Matérn correlation matrix based on the geographic centroids of each grid-cell as a random effect with the R package `r pack("spaMM")` [@Rousset2014]. 
We verified the significance of each term of the model by conducting likelihood ratio tests (LRTs) between the full model and models each with a single term of interest removed, in sequence. 
We also verified the significance of the full model relative to the null model (only including the Matérn correlation matrix) using LRTs.

@@

We analyzed bioregions using the framework of @Daru2020 and their R package `r pack("phyloregion")`. 
Briefly, this involves first calculating beta diversity (change in composition between communities) using either species names (taxonomic bioregions) or phylogenetic distances (phylogenetic bioregions). 
For taxonomic bioregions, we calculated the turnover in species between sites due to species replacement [@Baselga2012] with the Sørensen index. 
For phylogenetic bioregions, we calculated phylogenetic dissimilarities based on the PhyloSor index [@Bryant2008]. 
The distances are used to construct a dendrogram, which is then split into into *k* bioregions. 
The optimal value of *k* cannot be known *a-priori*.
Previous vegetation zone schemes in Japan have typically included 5--6 zones [@Shimizu2014]; to enable approximate comparison with these, we tested values of *k* from 1 to 10, and selected the optimal value using the `optimal_phyloregion` function in the R package `r pack("phyloregion")`.

@@

The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families. 
The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* [families follow @PteridophytePhylogenyGroupI2016]. 
The age of the fern clade is estimated to be `r phylogeny_age` ma.

@@

Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan (`r figure("rand-diversity")`a). 
A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Kii Peninsula). 
Grid-cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu (`r figure("rand-diversity")`b). 
Grid-cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido. 
Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`c, d).

@@

Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively. 
The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu), bioregion 3 (Ryukyu Islands, excluding Yakushima), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`a; `r s_figure("dendrogram")`a). 
Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion, and bioregion 1 extended slightly further south (`r figure("bioregions")`b; `r s_figure("dendrogram")`b). 
The other much smaller bioregions ("singletons" including only a single grid-cell each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.

@@

The overall range of Japanese deer in Japan has increased from `r deer_overall[["1978"]]$area` (`r deer_overall[["1978"]]$percent`) in 1978 to `r deer_overall[["2003"]]$area` (`r deer_overall[["2003"]]$percent`) in 2003, and may be as high as `r deer_overall$estimated$area` (`r deer_overall$estimated$percent`) according to model estimates.
The percentage of area occupied by deer in grid-cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
Cells with high FD, PD, or PE tended to have less deer occupation compared to the baseline rate (`r figure("conserv-status")`b).
However, cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.

@@

The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`a, `r s_figure("raw-lat")`a) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants [@Fukaya2020]. 
Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`a).
Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018; @Qian2021] to be important drivers of spatial fern diversity.
A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or sub-tropical areas [@Kessler2011a]. 
Hypotheses to explain such a pattern include purely geometrical constraints as well as biological hypothesis positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a]. 
Another reason for the greater richness in southern Honshu compared to smaller southern islands is that southern Honshu harbors high elevational variation that may support a wider range of taxa than is present in the southern islands.

The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`a, c). 
This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.). 
The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis, and suggests that in ferns, phylogeny can be used as a reasonable stand-in for functional diversity. 
Given their high phylogenetic diversity, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism; indeed, the vast majority of cells with significant levels of PE, as well as paleoendemic cells, are located in the southern islands. 
This is clearly due to the small land area of these islands, which are the only areas with a subtropical climate in Japan.

Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half of Honshu, and particularly on Kyushu (`r figure("rand-diversity")`b). 
The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a). The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`b) and *Dryopteris* (`r s_figure("genus-map")`d).

Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (`r figure("modelfit")`g), and to a slightly lesser extent SES of PD (`r figure("modelfit")`f).
Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in southern Honshu (`r s_figure("genus-map")`).
This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns. 
We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), a more direct measure of genetic diversity (`r s_figure("modelfit-not-ult")`b).
While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`). 
Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca. 34.8 °N latitude and increases with temperature in Japan (their study was limited to Honshu and did not include the southern islands). 
@Tanaka2014 asserted that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid. 
As fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses.

@@

The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`). 
Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014]. 
Furthermore, the border between bioregions 1 and 2 corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he proposed based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
This suggests that ferns are useful as bioindicators in Japan. 
Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from the rest of Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions. 
The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits bioregions 2 and 3, there is some overlap between them on the Amami-Oshima Isl. (`r figure("bioregions")`b). 
The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic bioregion 3 (mostly the Ryukyu islands; wetter) from bioregion 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).

@@

Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`b).
This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, where deer expansion has been intense (`r s_figure("protected-map")`a).
This result highlights the importance of using multiple metrics of biodiversity for establishing conservation priorities.

@@

Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_biogeo. 
A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_biogeo.
Data files needed to run the analysis and selected results files are available from Dryad at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.

>>>>

Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because the necessary data (DNA sequences and georeferenced occurrence records) and analytic tools have only become available relatively recently. 
These datasets and tools now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism. 
Better ways to characterize biological regions ("bioregions") are also available [e.g., @Laffan2016; @White2019; @Daru2020a], rather than relying on ad-hoc characterizations.
Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide powerful insights into the processes structuring biodiversity and suggest conservation priorities.

@@

A necessary step in any analysis of biodiversity is to set the grain size used to accurately define communities. 
Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. 
We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species). 
At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). 
We calculated sampling redundancy [1 - richness/number of specimens; @Garcillan2003] to quantify adequacy of sampling. 
Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
We filtered out any cells with redundancy < 0.1 and used this size for subsequent analyses.

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")` [@R-raster], and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`). 
We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by (`r figure("japan-map")`).

@@

Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]. 
However, using this dataset alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils available for molecular dating. 
Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree. 
To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa). 
We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with 1,000 ultrafast bootstrap replicates [@Minh2013].
Molecular dating was conducted with `r software("treePL")` v1.0 [@Smith2012] using `r n_fossil_points` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020]. 
The root of the tree (most recent common ancestor of all land plants) was fixed to 475 mya [@Testo2016a].
We trimmed the resulting ultrametric tree to only Japanese taxa.
treePL requires branch lengths to be > 0, and it sets extremely small branches to an arbitrary minimum positive length.
This may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* dataset.
We converted any clades in the dated tree consisting of identical sequences to polytomies with zero branch length between OTUs.
This tree was used for all subsequent analysis unless indicated otherwise.

@@

We tested for phylogenetic signal in continuous traits using two alternative metrics, Blomberg's *K* [@Blomberg2003] and Pagel's lambda [@Pagel1999]. 
*K* describes the ratio between the observed variance in a trait vs. the amount of variance expected under Brownian Motion (BM); lambda is a scaling parameter between 0 and 1 that transforms the phylogeny such that trait values fit those most closely expected under BM. 
Both lambda and *K* = 1 under BM; relatively lower values indicate less signal (overdispersion) and higher values indicate more signal (clustering). 
We measured *K* and lambda using the `r func("phylosig")` function in the R package `r pack("phytools")` [@Revell2012], which also calculates significance based on a randomization test for *K* and a likelihood ratio test for lambda. 
We tested for phylogenetic signal in qualitative (coded as binary) traits using Fritz and Purvis' *D* [@Fritz2010] with the `r func("phylo.d")` function in the R package `r pack("caper")` [@Orme2018], which also assesses significance by comparing the observed value to a null distribution of random values and a simulated distribution of values expected under BM.
These methods for phylogenetic signal cannot use a tree with zero-length polytomies, so we used the original dated tree from treePL.

@@

We constructed a linear mixed model for each biodiversity metric dependent on the environmental variables ("environmental models"). 
Initial inspection of the relationship between each biodiversity metric and temperature showed a hump-shaped pattern (Appendix S2), so we also included a quadratic term for temperature only. 
Richness was fit with the negative binomial response family; others were Gaussian. 
We fit models including a Matérn correlation matrix based on the geographic centroids of each grid-cell as a random effect with the R package `r pack("spaMM")` [@Rousset2014]. 
We verified the significance of each term of the model by conducting likelihood ratio tests (LRTs) between the full model and models each with a single term of interest removed, in sequence. 
We also verified the significance of the full model relative to the null model (only including the Matérn correlation matrix) using LRTs.

@@

We analyzed bioregions using the framework of @Daru2020 and their R package `r pack("phyloregion")`. 
Briefly, this involves first calculating beta diversity (change in composition between communities) using either species names (taxonomic bioregions) or phylogenetic distances (phylogenetic bioregions). 
For taxonomic bioregions, we calculated the turnover in species between sites due to species replacement [@Baselga2012] with the Sørensen index. 
For phylogenetic bioregions, we calculated phylogenetic dissimilarities based on the PhyloSor index [@Bryant2008]. 
The distances are used to construct a dendrogram, which is then split into into *k* bioregions. 
The optimal value of *k* cannot be known *a-priori*.
Previous vegetation zone schemes in Japan have typically included 5--6 zones [@Shimizu2014]; to enable approximate comparison with these, we tested values of *k* from 1 to 10, and selected the optimal value using the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")`.

@@

The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families. 
The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* [families follow @PteridophytePhylogenyGroupI2016]. 
The age of the fern clade is estimated to be `r phylogeny_age` ma.

@@

Grid-cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`a). 
A smaller number of cells had significantly high PD, and these were located mostly on small southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Kii Peninsula). 
Grid-cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`b). 
Grid-cells with significantly high RPD are located both in the small southern islands and northern Honshu and Hokkaido. 
Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`c, d).

@@

Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively. 
The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`a; `r s_figure("dendrogram")`a). 
Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 extended slightly further south (`r figure("bioregions")`b; `r s_figure("dendrogram")`b). 
The other much smaller bioregions ("singletons" including only a single grid-cell each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.

@@

The range of Japanese deer in Japan increased from `r deer_overall[["1978"]]$area` (`r deer_overall[["1978"]]$percent`) in 1978 to `r deer_overall[["2003"]]$area` (`r deer_overall[["2003"]]$percent`) in 2003, and may be as high as `r deer_overall$estimated$area` (`r deer_overall$estimated$percent`) according to model estimates.
The percentage of area occupied by deer in grid-cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
Cells with high FD, PD, or PE tended to have less deer occupation compared to the baseline rate (`r figure("conserv-status")`b).
However, cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.

@@

The latitudinal hump-shaped pattern of species richness observed in ferns here (Figs. `r figure_num("raw-diversity")`a, `r s_figure_num("raw-lat")`a) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants in Japan [@Fukaya2020]. 
Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`a).
Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018; @Qian2021] to be important drivers of spatial fern diversity.
A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or sub-tropical areas [@Kessler2011a; @Suissa2021]. 
Hypotheses to explain such a pattern include purely geometrical constraints as well as biological hypothesis positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a]. 
Another reason for the greater richness on the main islands compared to southern islands in Japan is that the main islands harbor higher elevational variation that may support a wider range of taxa than is present in the southern islands.

The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`a, c). 
This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.). 
The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Tables `r s_table_num("phy-sig")`, `r s_table_num("phy-sig-binary")`), and suggests that in ferns, phylogeny can be used as a reasonable stand-in for functional diversity. 
Given their high phylogenetic diversity, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism; indeed, the vast majority of cells with significant levels of PE, as well as paleoendemic cells, are located in the southern islands (Figs. `r figure_num("endemism")`, `r s_figure_num("endemism-restricted")`). 
This is clearly due to the small land area of these islands, which are the only areas with a subtropical climate in Japan.

Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and the southern islands, and the predominance of short branches in southern Honshu, Shikoku, and Kyushu (`r figure("rand-diversity")`b). 
The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a). The preponderance of short branches in southern Honshu, Shikoku, and Kyushu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`b) and *Dryopteris* (`r s_figure("genus-map")`d).

Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (Figs. `r figure_num("coeff")`g, `r figure_num("modelfit")`g), and to a slightly lesser extent SES of PD (Figs. `r figure_num("coeff")`f, `r figure_num("modelfit")`f).
Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in the southern main islands (`r s_figure("genus-map")`).
This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns. 
We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), a more direct measure of genetic diversity (`r s_figure("modelfit-not-ult")`).
While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`). 
Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca. 34.8 °N latitude and increases with temperature in Japan (their study was limited to the main islands and did not include the smaller southern islands). 
@Tanaka2014 asserted that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid. 
As fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses to our knowledge.

@@

The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`). 
Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014]. 
Furthermore, the border between bioregions 1 and 2 corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he proposed based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
This suggests that ferns are useful as bioindicators in Japan. 
Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions. 
The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits phylogenetic bioregions 2 and 3, there is some overlap between them on Yakushima and Amami Isl. (`r figure("bioregions")`b). 
The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind [@Tryon1970].
Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic bioregion 3 (mostly the Ryukyu islands; wetter) from bioregion 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).

@@

Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`b).
This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (`r s_figure("protected-map")`a).
This result highlights the importance of using multiple metrics of biodiversity for establishing conservation priorities.

@@

Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_spatial_phy. 
A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_spatial_phy.
Data files needed to run the analysis and selected results files are available from Dryad at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.
